### 2024-05-10 code/draft-script.Rmd
run test script for some mcda

---
title: "Untitled"
author: "Emily Zhou"
date: "2023-11-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load packages, message=FALSE, warning=FALSE}
library(tidyverse)
library(sf)
library(RSocrata)
library(viridis)
library(spatstat.explore)
library(raster)
library(spdep)
library(FNN)
library(grid)
library(gridExtra)
library(knitr)
library(kableExtra)
library(tidycensus)
library(classInt)   
library(svDialogs)
library(patchwork)
library(tidycensus)
library(SpatialAcc)
library(tmap)
library(here)
library(svDialogs)
source("https://raw.githubusercontent.com/urbanSpatial/Public-Policy-Analytics-Landing/master/functions.r")
```

```{r}
census_api_key(dlgInput("Enter a Census API Key", 
  Sys.getenv("CENSUS_API_KEY"))$res,
  overwrite = TRUE)
```

# Prepare Analysis Layer

```{r philly fishnet, message=FALSE, warning=FALSE}

philly <- st_read(here("data", "raw", "city_boundary.geojson")) %>% st_transform('ESRI:102728')
fishnet <- st_make_grid(philly,
               cellsize = 1640, # 500 meters
               square = TRUE) %>%
  .[philly] %>%           
  st_sf() %>%
  mutate(uniqueID = 1:n())
```

```{r remove water, message=FALSE, warning=FALSE}

water <- st_read(here("data", "raw", "water_features.geojson")) %>% st_transform('ESRI:102728')
inWater <- water %>% 
  st_buffer(dist = 330) %>% 
  st_intersection(st_centroid(fishnet), .) %>% 
  st_drop_geometry() %>% 
  dplyr::select(uniqueID)

fishnet_nowater<- fishnet %>%
  filter(!(uniqueID %in% inWater$uniqueID))

```

```{r grids to analysis}

ggplot() +
  geom_sf(data=fishnet, color="black", fill="blue") +
  geom_sf(data=fishnet_nowater, color="black", fill="white") +
  labs(title = "Fishnet of Philly") +
  theme(axis.text.x=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks =element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        plot.subtitle = element_text(size = 9,face = "italic"),
        plot.title = element_text(size = 12, face = "bold"),
        panel.background = element_blank(),
        panel.border = element_rect(colour = "grey", fill=NA, linewidth=0.8)
        )
```

# Load Criteria

```{r population density and driving pop}

net_centroid <- st_centroid(fishnet_nowater)
acs_variable_list.2020 <- load_variables(2020, #year
                                         "acs5", #five year ACS estimates
                                        cache = TRUE)

options(timeout=1000) 
philly20 <- get_acs(geography = "tract", 
          variables = c(
            "B01001_001E", # total population
            "B01001_010E", # male 22-24
            "B01001_011E",
            "B01001_012E",
            "B01001_013E",
            "B01001_014E",
            "B01001_015E",
            "B01001_016E",
            "B01001_017E",
            "B01001_018E",
            "B01001_019E", # male 62-64
            "B01001_034E",
            "B01001_035E",
            "B01001_036E",
            "B01001_037E",
            "B01001_038E",
            "B01001_039E",
            "B01001_040E",
            "B01001_041E",
            "B01001_042E",
            "B01001_043E" #female 62-64
            ), 
          year=2020, state="PA", county="Philadelphia", 
          geometry=TRUE, output="wide") %>%
  st_transform('ESRI:102728')

philly20 <- philly20 %>% 
  mutate(popden = B01001_001E / (st_area(geometry)*9.2903e-8),
         density = as.integer(gsub("\\[.*\\]", "", popden)),
         pop22_64 = B01001_010E + B01001_011E + B01001_012E+ B01001_013E+ B01001_014E+ B01001_015E+ B01001_016E+ B01001_017E+ B01001_018E+ B01001_019E + B01001_034E + B01001_035E + B01001_036E + B01001_037E + B01001_038E + B01001_039E + B01001_040E + B01001_041E + B01001_042E + B01001_043E,
         driving_pop = pop22_64 / B01001_001E,
         totalpop = B01001_001E) %>% 
  dplyr::select(GEOID, totalpop, density, driving_pop)

philly20$driving_pop[is.nan(philly20$driving_pop)] <- 0

```

```{r existing EVSE}

evse <- read.csv(here("data", "raw", "existing_evse.csv"))
evse <- evse %>% 
  filter(City == "Philadelphia") %>% 
  dplyr::select(Latitude, Longitude, EV.Level2.EVSE.Num) %>% 
  filter(is.na(Latitude) == FALSE & is.na(Longitude) == FALSE) %>% 
  st_as_sf(., coords = c("Longitude", "Latitude"), crs = 4326) %>% 
  st_transform('ESRI:102728') %>% 
  mutate(num = ifelse(is.na(EV.Level2.EVSE.Num), 1, EV.Level2.EVSE.Num))

fishnet_nowater <- fishnet_nowater %>%
  left_join(net_centroid %>% 
              mutate(evse.nn = nn_function(st_coordinates(net_centroid), 
                                           st_coordinates(evse), 2)*0.3048) %>% 
              st_drop_geometry(), 
            by = "uniqueID")

```

```{r spatial accessibility}

tracts.coords <- st_coordinates(st_centroid(philly20))
evse.coords <- st_coordinates(evse)
dist.matrix <- distance(tracts.coords, evse.coords, type = "euclidean")

TSFCA <- ac(p = philly20$totalpop, 
           n = evse$num, 
           D = dist.matrix, d0 = 2000, family = "2SFCA")

philly20 <- philly20 %>% 
  mutate(TSFCA = TSFCA)

fishnet_nowater <- fishnet_nowater %>%
  left_join(net_centroid %>% st_intersection(philly20) %>% st_drop_geometry(), by = "uniqueID")

```

```{r map accessibility}

tm_shape(philly20, unit = "ft") +
  tm_polygons(col = "TSFCA", style = "jenks",palette = "viridis", 
              border.alpha = 0, title = "2SFCA") +
    tm_layout(main.title = "Spatial Accessibility of EVSE in Philadelphia County",
            main.title.size = 0.95, frame = FALSE,
            legend.outside = TRUE, legend.outside.position = "right") + 
   tm_credits("Using Two-Step Floating Catchment Area (2000ft) Calculation",
             position = c("RIGHT", "BOTTOM"), size = 0.5)

#tmap_save(filename = "accessibility_by_tract.png")
```

https://afdc.energy.gov/fuels/electricity_locations.html#/analyze?fuel=ELEC&region=US-PA

```{r registered EV}

reg_ev <- st_read(here("data", "raw", "reg_ev.geojson"))

Zip_Code <- as.character(c(
  19120, 19124, 19111, 19143, 19149, 19134, 19140, 19148, 19104, 19144,
  19145, 19131, 19139, 19146, 19147, 19115, 19136, 19128, 19135, 19121,
  19154, 19141, 19132, 19152, 19114, 19116, 19151, 19138, 19142, 19119,
  19130, 19125, 19133, 19103, 19150, 19122, 19126, 19123, 19107, 19106,
  19153, 19129, 19118, 19137, 19127, 19102, 19108, 19109, 19176, 19112,
  19160, 19155, 19162, 19161, 19171, 19170, 19173, 19172, 19175, 19178,
  19177, 19181, 19179, 19183, 19182, 19185, 19184, 19188, 19187, 19192,
  19191, 19194, 19193, 19196, 19195, 19244, 19197, 19255, 19019, 19093,
  19092, 19101, 19099, 19105, 19110, 19190
))

reg_ev <- reg_ev %>% 
  filter(zip %in% Zip_Code) %>% 
  dplyr::select(elec_cnt) %>% 
  st_transform('ESRI:102728')

fishnet_nowater <- fishnet_nowater %>%
  left_join(net_centroid %>% st_intersection(reg_ev) %>% st_drop_geometry(), by = "uniqueID")

```

```{r map registered EV}

tm_shape(reg_ev, unit = "ft") +
  tm_polygons(col = "elec_cnt", style = "jenks",palette = "viridis", 
              border.alpha = 0, title = "EV") +
    tm_layout(main.title = "Registered EV by Zip Code in Philadelphia County",
            main.title.size = 0.95, frame = FALSE,
            legend.outside = TRUE, legend.outside.position = "right")

#tmap_save(filename = "reg_ev_zip.png")
```

```{r parking lots}

parking <- st_read(here("data", "raw", "osm_features", "parking.geojson"))

parking <- parking %>% 
  filter(!access %in% c("no", "permissive", "permit", "private")) %>% 
  st_transform('ESRI:102728')

fishnet_nowater <- fishnet_nowater %>%
  left_join(net_centroid %>% 
              mutate(parking.nn = nn_function(st_coordinates(net_centroid), 
                                           st_coordinates(parking), 2)*0.3048) %>% 
              st_drop_geometry(), 
            by = "uniqueID")

```

```{r add police and fire stations}

police_fire <- st_read(here("data", "raw", "osm_features", "police_fire.geojson")) %>% st_transform('ESRI:102728') %>% st_centroid()

fishnet_nowater <- fishnet_nowater %>%
  left_join(net_centroid %>% 
              mutate(policefire.nn = nn_function(st_coordinates(net_centroid), 
                                           st_coordinates(police_fire), 1)*0.3048) %>% 
              st_drop_geometry(), 
            by = "uniqueID") 
```

```{r add zoning rules}

zoning <- st_read(here("data", "raw", "osm_features", "zoning.shp"))

zoning <- zoning %>% st_transform('ESRI:102728')

fishnet_nowater <- fishnet_nowater %>%
  left_join(net_centroid %>% st_buffer(dist = 100) %>% st_intersection(zoning) %>% st_drop_geometry(), by = "uniqueID")
```

```{r}

tst <-  fishnet_nowater %>% 
  #filter(is.na(GEOID) == FALSE) %>% 
  mutate(zoning = case_when( ZONINGGROU == "Commercial/Commercial Mixed-Use" ~ "Commercial",
                             ZONINGGROU == "Industrial/Industrial Mixed-Use" ~ "Industrial",
                             ZONINGGROU == "Residential/Residential Mixed-Use" ~ "Residential",
                             ZONINGGROU == "Special Purpose" ~ "Special",
                             TRUE ~ "Undefined")) %>% 
  mutate(zone_score = case_when( zoning == "Commercial" ~ 0.4,
                             zoning == "Industrial" ~ 0.15,
                             zoning == "Residential" ~ 0.3,
                             TRUE ~ 0.15)) %>% 
  mutate(elec_cnt = ifelse(is.na(elec_cnt), 0, elec_cnt)) %>% 
  dplyr::select(-featCount, -ZONINGGROU) %>% 
  mutate(TSFCA = ifelse(is.infinite(TSFCA), 0, TSFCA))

```

## Pairwise comparison matrix

```{r}

pw_df <- data.frame(
  popden = c(1,8,3,5,5,4,2,2),
  acc = c(1/8, 1, 1/4, 1/3, 1/4, 1/5, 1/2, 1/2),
  driving = c(1/3, 4, 1, 1/2, 1/5, 1/2, 1/4, 1/4),
  evse = c(1/5, 3,2,1, 1/2, 1/4, 1/5, 1/5),
  reg_ev = c(1/5, 4, 5, 2, 1, 1/5, 1/5, 1/5),
  parking = c(1/4, 5, 2, 4, 5, 1, 1/3, 1/4),
  police = c(1/2, 2, 4, 5, 5, 3, 1, 1/2),
  zoning = c(1/2, 2, 4, 5, 5, 4, 2, 1),
 row.names = c("popden", "acc", "driving", "evse", "reg_ev", "parking", "police", "zoning")
)

pw_df <- pw_df %>% 
  mutate(unweight = (popden * acc * driving * evse * reg_ev * parking * police * zoning)^0.125) %>% 
  mutate(weighted = unweight / sum(unweight))

```

```{r}

pw_df_2 <- data.frame(
  popden = c(1, 7, 5, 3, 2),
  acc = c(1/7, 1, 1/2, 1/3, 1/5),
  reg_ev = c(1/5, 2, 1, 1/5, 1/3),
  parking = c(1/3, 3, 5, 1, 1/3),
  zoning = c(1/2, 5, 3, 3, 1),
  row.names = c("popden", "acc", "reg_ev", "parking", "zoning")
)

pw_df_2 <- pw_df_2 %>% 
  mutate(unweight = (popden * acc * reg_ev * parking * zoning)^0.2) %>% 
  mutate(weighted = unweight / sum(unweight))

```

## Consistency Check

```{r}

A1 <- pw_df_2%>% 
  dplyr::select(-c(unweight, weighted)) %>% 
  as.matrix(rownames.force = NA)

A2 <- pw_df_2 %>% 
  dplyr::select( weighted) %>% 
  as.matrix()

A3 <- A1 %*%  A2

A4 <- A3 / A2
A4 <- as.data.frame(A4)

((mean(A4$weighted) - 5) / 4) / 1.11
```

## Weighted sum method

## TOPSIS method

```{r scale}

scale_values <- function(x){(x-min(x))/(max(x)-min(x))}

tst <- tst %>% 
  mutate(scl_density = scale_values(density),
         scl_acc = scale_values(TSFCA),
         scl_drivingpop = scale_values(driving_pop),
         scl_evse = scale_values(evse.nn),
         scl_elec = scale_values(elec_cnt),
         scl_parking = scale_values(parking.nn),
         scl_policefire = scale_values(policefire.nn)
         ) %>% # now inverse some scales
  mutate(scl_evse_re = 0 - scl_evse + 1,
         scl_acc_re = 0 - scl_acc + 1, 
         scl_parking_re = 0 - scl_parking + 1,
         scl_policefire_re = 0 - scl_policefire + 1)

```

```{r}

tst <- tst %>% 
  mutate(score = 0.15*scl_density + 0.2*scl_evse_re + 0.2*scl_elec + 0.1*scl_drivingpop + 0.1*scl_parking_re + 0.05*scl_policefire_re + 0.05*zone_score + 0.15*scl_acc_re)

```

```{r}

neighborhood <- st_read("/Users/emzhou/Documents/Fall2023/CPLN-501-Quant/PhillyPlanning_Neighborhoods/PhillyPlanning_Neighborhoods.shp") %>% st_transform('ESRI:102728')

tst <- tst %>%
  left_join(net_centroid %>% 
               st_intersection(neighborhood %>% dplyr::select(NAME)) %>% 
              st_drop_geometry(), by = "uniqueID") %>% 
  mutate(NAME = ifelse(is.na(NAME), "NOT APPLICABLE", NAME))

```

```{r}

output_file <- paste0("fishnet", ".shp")
st_write(fishnet, output_file, append = FALSE)

output_file2 <- paste0("accessibility", ".shp")
st_write(tst, output_file2, append = FALSE)

```

```{r}
ggplot()+
  geom_sf(data=fishnet, color="black", fill="white") +
  geom_sf(data=tst, color="black", aes(fill=score))
```

```{r}

library(leaflet)
summary(tst$score)

```

```{r}

acme <- read.csv("/Users/emzhou/Desktop/sc-final-proj/acme.csv") %>% 
  st_as_sf(., coords = c("Longitude", "Latitude"), crs = 4326) %>% 
  mutate(longitude= unlist(map(geometry, 1)),
         latitude = unlist(map(geometry, 2)))

```

```{r}

pal <- colorBin(palette = "viridis", # change the color palette here
                domain = tst$score, # change the parameter name after $
                bins = c(0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9)) # change the bins here

```

```{r}

labs <- paste0("<strong>", tst$NAME,
               "</strong><br/>EVSE Suitability: ", tst$score, 
               "</strong><br/>Population Density (sqkm): ", tst$density,
               "</strong><br/>Accessibility Score: ", tst$TSFCA,
               "</strong><br/>Driving Population %: ", tst$driving_pop,
               "</strong><br/> Registered EV: ", tst$elec_cnt,
               "</strong><br/>Avg Distance to EVSE: ", tst$evse.nn,
               "</strong><br/>Avg Distance to Parking: ", tst$parking.nn,
               "</strong><br/>Avg Distance to Police & Fire: ", tst$evse.nn
               )

labs_acme <- paste0("</strong><br/>ACME Address ", acme$Address)
```

```{r show map}

leaflet <- tst %>%
  st_transform("EPSG:4326") %>% 
  leaflet() %>%
  setView(lng = -75.1652, lat = 40.0126, zoom = 10.5) %>%
  addProviderTiles(providers$CartoDB.Positron) %>% 
  addPolygons(weight = 1,
              fillColor = ~pal(score), # change the name of the parameter in the bracket
              popup = labs,
              fillOpacity = 1) %>%
  addCircleMarkers(
    data = acme,  # Replace with your actual point data
    lng = ~longitude,  # Replace with the appropriate variable in your point dataset for longitude
    lat = ~latitude,  # Replace with the appropriate variable in your point dataset for latitude
    weight = 0.5,
    radius = 4,
    fillColor = "red",  # Set the fill color for the markers
    fillOpacity = 1,
    popup = ~labs_acme  # Replace with the appropriate variable in your point dataset for popup text
  ) %>% 
  addLegend(position = "bottomright",
            pal = pal,
            values = ~score, # change the name of the parameter after ~
            title = "EVSE Suitability Score") # change the name of the legend

leaflet
```

```{r}

mapshot(leaflet, url = "evse-suitability.html")

```

	Population density	Spatial accessibility	Registered EV	Distance to Parking	zoning
Population density	1	1/7	1/5	1/3	1/2
Spatial accessibility	7	1	2	3	5
Registered EV	5	1/2	1	5	3
Distance to Parking	3	1/3	1/5	1	3
zoning	2	1/5	1/3	1/3	1