### 2024-05-22 code/test-mcda.Rmd
testing the promethee method

---
title: "test-mcda"
author: "Emily Zhou"
date: "2024-05-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Testing The MCDA Package

```{r}
Rglpk, glpkAPI, RColorBrewer, combinat, triangle, plyr,
ggplot2
```

## Testing the AHP Approach 
```{r}
alts <- c("Corsa","Clio","Fiesta","Sandero")

style <- matrix(c(1.0, 1/4, 4.0, 1/6,
4.0, 1.0, 4.0, 1/4,
1/4, 1/4, 1.0, 1/5,
6.0, 4.0, 5.0, 1.0),
nrow=length(alts), ncol=length(alts), byrow=TRUE,
dimnames=list(alts,alts))

reliability <- matrix(c(1.0, 2.0, 5.0, 1.0,
1/2, 1.0, 3.0, 2.0,
1/5, 1/3, 1.0, 1/4,
1.0, 1/2, 4.0, 1.0),
nrow=length(alts), ncol=length(alts), byrow=TRUE,
dimnames=list(alts,alts))

fuel <- matrix(c(1.0, 2.0, 4.0, 1.0,
0.5, 1.0, 3.0, 2.0,
1/4, 1/3, 1.0, 1/5,
1.0, 1/2, 5.0, 1.0),
nrow=length(alts), ncol=length(alts), byrow=TRUE,
dimnames=list(alts,alts))
```

```{r}

alternativesPairwiseComparisonsList <- list(style = style,
reliability = reliability,
fuel = fuel)

```

```{r}
crit <- c("style","reliability","fuel")
criteriaWeightsPairwiseComparisons <- matrix(c(1.0, 1/2, 3.0,
2.0, 1.0, 4.0,
1/3, 1/4, 1.0),
nrow=length(crit),
ncol=length(crit),
dimnames=list(crit,crit))
```

```{r}
AHP(criteriaWeightsPairwiseComparisons, alternativesPairwiseComparisonsList)
```

## Testing the TOPSIS Approach

```{r}

performanceTable <- matrix(c(5490,51.4,8.5,285,6500,70.6,7,
                             288,6489,54.3,7.5,290),
                           nrow=3,
                           ncol=4,
                           byrow=TRUE)

row.names(performanceTable) <- c("Corsa","Clio","Fiesta")
colnames(performanceTable) <- c("Purchase Price","Economy",
                                "Aesthetics","Boot Capacity")

weights <- c(0.35,0.25,0.25,0.15)
criteriaMinMax <- c("min", "max", "max", "max")

# positiveIdealSolutions <- c(0.179573776, 0.171636015, 0.159499658, 0.087302767)
# negativeIdealSolutions <- c(0.212610118, 0.124958799, 0.131352659, 0.085797547)
names(weights) <- colnames(performanceTable)

names(criteriaMinMax) <- colnames(performanceTable)
# names(positiveIdealSolutions) <- colnames(performanceTable)
# names(negativeIdealSolutions) <- colnames(performanceTable)
```

```{r}
overall1 <- TOPSIS(performanceTable, weights, criteriaMinMax)
overall1 # returns a vector containing the TOPSIS score for each alternative
```

## Testing the PROMETHEE Approach

The function returns three matrices: The first one contains the preference relations between the
alternatives, the second one contains the indifference relations between the alternatives and the
third one contains the incomparability relations between the alternatives.

```{r}

performanceTable <- rbind(
c(1,10,1),
c(4,20,2),
c(2,20,0),
c(6,40,0),
c(30,30,3))
rownames(performanceTable) <- c("RER","METRO1","METRO2","BUS","TAXI")
colnames(performanceTable) <- c("Price","Time","Comfort")
```

```{r}
# The preference functions
preferenceFunction<-c("Gaussian","Level","V-shape-Indiff")
```

```{r}
#Preference threshold
preferenceThreshold<-c(5,15,3)
names(preferenceThreshold)<-colnames(performanceTable)
```

```{r}
#Indifference threshold
indifferenceThreshold<-c(3,11,1)
names(indifferenceThreshold)<-colnames(performanceTable)

#Parameter of the Gaussian preference function
gaussParameter<-c(4,0,0)
names(gaussParameter)<-colnames(performanceTable)

#weights
criteriaWeights<-c(0.2,0.3,0.5)
names(criteriaWeights)<-colnames(performanceTable)

# criteria to minimize or maximize
criteriaMinMax<-c("min","min","max")
names(criteriaMinMax)<-colnames(performanceTable)

PROMETHEEOutrankingFlows(performanceTable, preferenceFunction,preferenceThreshold,
indifferenceThreshold,gaussParameter,criteriaWeights,criteriaMinMax)
```